---
title: "2 Cross-Lagged Structural Equation analysis"
author: "Tomoe Gusberti, Dr. Eng"
date: "2025-03-20"
output: pdf_document
always_allow_html: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lavaan)
#library(tidySEM)
#detach("package:lavaan.mi", unload = TRUE)
library(lavaan.mi)
#library(semTools)

library(ggplot2)
```

# PREPARE DATA
- depression levels- from imputed (multiple) data
- multimorbidity from original scores?
- inverted propensity weight - from ipwBalanced (imputation does not change it)
```{r}
load(file='./Data/Data_Imp20.RData')#imp20
```

```{r}
load(file='./Data/Data_ipwBalancedT3_wide.RData')#data_bal
data_bal<-data_bal%>%mutate(depression=ifelse(Depressao0Cat=='normal','non-Depressive',
                                      ifelse(Depressao0Cat%in%c('mild',
                                      'moderate','severe'),'Depressive',NA)))%>%
    rename('Dep_EpiDoc1' := Depressao_T0, 'Dep_EpiDoc2':= Depressao_T1,
           'Dep_EpiDoc4':= Depressao_T3,
           'Age':=Idade_T0)

```
## merging and creating as mids
```{r}
library(miceadds)
#datlist2mids

data_mi=list()
for (i in 1:imp20$m){
  dt<-complete(imp20,i)
  dt<-left_join(data_bal%>%
                  select(ID,ipw,depression,MMorb_T0,MMorb_T1,MMorb_T3,MMorb0)
                ,dt) %>%
    mutate(
      MMorbidity=ifelse(MMorb0>=2,'>=2 diseases','<2 diseases'),
      EducationalLevel=ifelse(Escolaridade>2,
                              '<=9 years','>9years')
      )%>%
    #T0=EpiDoc1, T1=Epipdoc2, T3=EpiDoc4
    rename('Dep_EpiDoc1' := Pontuacao_Depressao_T0, 'Dep_EpiDoc2':= Pontuacao_Depressao_T1,
           'Dep_EpiDoc4':= Pontuacao_Depressao_T3,
           'MM_EpiDoc1':=MMorb_T0,
           'MM_EpiDoc2':=MMorb_T1,
           'MM_EpiDoc4':=MMorb_T3,
           'Age':=Idade_T0,
           'Education':=Escolaridade)%>%mutate(
             Sex=ifelse(SEXO==1,'Male','Female')
           )
  # prep interaction terms
  dt<-dt%>%mutate(
    MM2_ed=scale(Education*MM_EpiDoc2)%>%as.numeric(),
    MM1_ed=scale(Education*MM_EpiDoc1)%>%as.numeric(),
    d2_ed=scale(Education*Dep_EpiDoc2)%>%as.numeric(),
    d1_ed=scale(Education*Dep_EpiDoc1)%>%as.numeric(),
    age_ed=scale(Education*Age)%>%as.numeric()
  )
  data_mi[[i]]<-dt
}

data.mi=datlist2mids(data_mi)
save(data.mi,file='./Data/Data_Imp20_forSEM.RData')

```

# Lagged SEM model


## ondas 0 e 1
```{r}
model01<-'
Dep_EpiDoc2~MM_EpiDoc1+Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1+Dep_EpiDoc1
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
'
res<-lavaan.mi::sem.mi(data=data.mi,std.ov=TRUE,
         model=model01,# missing='pairwise',
         estimator='MLR',likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',
         orthogonal=TRUE
         )
summary(res,fit.meas=TRUE)
```

## ondas 1, 2 e 3

```{r}
model013<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4
'
res<-lavaan.mi::sem.mi(data=data.mi,std.ov=TRUE,
         model=model013,# missing='pairwise',
         estimator='MLR',likelihood = "wishart", #missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',
         orthogonal=TRUE
         )
summary(res,fit.meas=TRUE)
```

# adding age
```{r}
modelAll0<-'
# evolução
Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
#Dep_EpiDoc4~MM_EpiDoc2
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~dm1*MM_EpiDoc1
MM_EpiDoc2~md1*Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~d3*Age
MM_EpiDoc4~m3*Age
Dep_EpiDoc2~d1*Age
MM_EpiDoc2~m1*Age
Dep_EpiDoc1~d0*Age
MM_EpiDoc1~m0*Age

age_dep_indirect:=d0*dd01*dd13+d0*dd03
age_dep_direct:=d3

dep0_direct:=dd03
dep0_indirect:=dd01*dd13
dep_total:=dd03+dd13*dd01

# hypotheses 3
age_dep1_direct:=d1
age_dep1_m_indirect:=m0*dm1

'
resAll0<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=modelAll0,# missing='pairwise',
         estimator='MLR',likelihood = "wishart", #missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',
         orthogonal=TRUE
         )
summary(resAll0,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
```{r}
source('./code/parTableToCSV.R')
parTableToCSV(resAll0,path='./Results/LaggedSEM_mi20/g_SEMED')
```
# adjusted and education
```{r}
modelAll0_adj<-'
# evolução
Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~mm13*MM_EpiDoc2+mm03*MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~mm01*MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~md13*MM_EpiDoc2+md03*MM_EpiDoc1
#Dep_EpiDoc4~md13*MM_EpiDoc2
MM_EpiDoc4~dm13*Dep_EpiDoc2+dm03*Dep_EpiDoc1
Dep_EpiDoc2~dm01*MM_EpiDoc1
MM_EpiDoc2~md01*Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~d3*Age
MM_EpiDoc4~m3*Age
Dep_EpiDoc2~d1*Age
MM_EpiDoc2~m1*Age
Dep_EpiDoc1~d0*Age
MM_EpiDoc1~m0*Age

# educação
Dep_EpiDoc4~de3*Education
MM_EpiDoc4~me3*Education
Dep_EpiDoc2~de1*Education
MM_EpiDoc2~me1*Education
Dep_EpiDoc1~de0*Education
MM_EpiDoc1~me0*Education

#age effects
age_dep_indirect:=d0*dd01*dd13+d0*dd03+d1*dd13
age_depM_indirect:=m0*md03+m1*md13+d0*dm01*md13+d0*dm01*md13
age_direct:=d3
age_total:=age_dep_indirect+age_depM_indirect


## indirect effect of educaiton on dep4
ed_dep_indirect:=de0*dd01*dd13+de1*dd13
ed_depM_indirect:=de1*dm01*md13
ed_direct:=de3

ed_total:=ed_dep_indirect+ed_depM_indirect

## cumulative effect d1 on dep4
dep0_direct:=dd03
dep0_indirect:=dd01*dd13+dd01*dd13
dep0_m_indirect:=dm01*md13+dm01*md13
dep_total:=dep0_direct+dep0_indirect+dep0_m_indirect

##indirect effect MM1 on dep4
m0_direct:=md03
m0_indirect:=md01*dd13+mm01*md13
m0_total:=m0_direct+m0_indirect


'
res_adj<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=modelAll0_adj,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart", missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',
         orthogonal=TRUE
         )
summary(res_adj,fit.meas=TRUE)
```
```{r}
parameterEstimates.mi(res_adj)
anova(res_adj)
```


```{r}
source('./code/parTableToCSV.R')
res<-parTableToCSV(res_adj,path=('./Results/LaggedSEM_mi20_final'))
res$partable
```

# Multi-group assessment
```{r}
modelfree<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc4~MM_EpiDoc2
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~Age
Dep_EpiDoc2~Age
MM_EpiDoc2~Age
Dep_EpiDoc1~Age
MM_EpiDoc1~Age

# educação
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'
```

```{r}

resFre_g<-lavaan.mi::sem.mi(data=data.mi,
         model=modelfree,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(resFre_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
## theoretical restrictions on age, auto-regressive  or, cross-lagged components
### restriction on residual cov
```{r}

model_covR<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~Age
Dep_EpiDoc1~Age

MM_EpiDoc4~Age
MM_EpiDoc2~Age
MM_EpiDoc1~Age

# educação
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'


res_covR_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covR,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covR_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g)
```

### restriction on age direct effect
- we free equality on cross-lagged effect... but if gender (and contextual issues matter, time point 4 would differ also... )
- equal among group (doubt/hypotheses under test here)

#### fully restrictive - on equality
```{r}

model_covAge_R<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~d2*Age
Dep_EpiDoc1~d1*Age

MM_EpiDoc4~m4*Age
MM_EpiDoc2~m2*Age
MM_EpiDoc1~m1*Age

# educação
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'


res_covAge_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covAge_R,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covAge_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g,res_covAge_g)
```
difference is significant, constraining to make the effect of age being the  same among groups leads to worsening of the fit.

```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_covAge_g,path=('./Results/LaggedSEM_mi20/gSEX_covAge'))
res$gPartable

```

#### freeing up MM
- we free equality on cross-lagged effect... but if gender (and contextual issues matter, time point 4 would differ also... )
- equal among group for dep (doubt/hypotheses under test here) 
```{r}

model_covAge_R<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1


# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~ad2*Age
Dep_EpiDoc1~ad1*Age

MM_EpiDoc4~c(am4a,am4b)*Age
MM_EpiDoc2~c(am2a,am2b)*Age
MM_EpiDoc1~c(am1a,am1b)*Age



# educação
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'


res_covAge_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covAge_R,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covAge_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g,res_covAge_g)
```
difference is still significant, constraining to make the effect of age being the  same among groups leads to worsening of the fit.

```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_covAge_g,path=('./Results/LaggedSEM_mi20/gSEX_covAge'))
res$gPartable

```


```{r}

model_covAge_R<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1


# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~ad2*Age
Dep_EpiDoc1~ad1*Age

MM_EpiDoc4~c(am4a,am4b)*Age
MM_EpiDoc2~c(am2a,am2b)*Age
MM_EpiDoc1~c(am1a,am1b)*Age



# educação
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'


res_covAge_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covAge_R,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covAge_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g,res_covAge_g)
```
difference is still significant, constraining to make the effect of age being the  same among groups leads to worsening of the fit.

```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_covAge_g,path=('./Results/LaggedSEM_mi20/gSEX_covAge'))
res$gPartable

```

### restriction on education 
#### restrictions on NO direct effect
```{r}

model_covAgeEd_R<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1


# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~ad2*Age
Dep_EpiDoc1~ad1*Age

MM_EpiDoc4~c(am4a,am4b)*Age
MM_EpiDoc2~c(am2a,am2b)*Age
MM_EpiDoc1~c(am1a,am1b)*Age



# educação
Dep_EpiDoc4~0*Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~0*Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education
'


res_covAgeEd_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covAgeEd_R,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covAgeEd_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g,res_covAge_g,res_covAgeEd_g)
```
difference is still significant, constraining to make the effect of age being the  same among groups leads to worsening of the fit.

```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_covAgeEd_g,path=('./Results/LaggedSEM_mi20/gSEX_covAgeEd'))
res$gPartable

```


#### restrictions on NO effect MM
```{r}

model_covAgeEd2_R<-'
# evolução
Dep_EpiDoc4~c(dd14a,dd14b)*Dep_EpiDoc1+c(dd24a,dd24b)*Dep_EpiDoc2
Dep_EpiDoc2~c(dd12a,dd12b)*Dep_EpiDoc1

MM_EpiDoc4~c(mm24a,mm24b)*MM_EpiDoc2+c(mm14a,mm14b)*MM_EpiDoc1
MM_EpiDoc2~c(mm12a,mm12b)*MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~c(md24a,md24b)*MM_EpiDoc2+c(md14a,md14b)*MM_EpiDoc1
Dep_EpiDoc2~c(md12a,md12b)*MM_EpiDoc1

MM_EpiDoc4~c(dm24a,dm24b)*Dep_EpiDoc2+c(dm14a,dm14b)*Dep_EpiDoc1
MM_EpiDoc2~c(dm12a,dm12b)*Dep_EpiDoc1


# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~ad2*Age
Dep_EpiDoc1~ad1*Age

MM_EpiDoc4~c(am4a,am4b)*Age
MM_EpiDoc2~c(am2a,am2b)*Age
MM_EpiDoc1~c(am1a,am1b)*Age



# educação
Dep_EpiDoc4~0*Education
Dep_EpiDoc2~c(ed2a,ed2b)*Education
Dep_EpiDoc1~c(ed1a,ed1b)*Education
MM_EpiDoc4~0*Education
MM_EpiDoc2~0*Education
MM_EpiDoc1~0*Education


# effect on multimorbidity


# effects on depression =============
##indirect effect age on dep4
ada:=ad1*dd14a+ad2*dd24a+ad1*dd12a*dd24a
adb:=ad1*dd14b+ad2*dd24b+ad1*dd12b*dd24b
amda:=am1a*md14a+am2a*md24a+am1a*mm12a*md24a+ad1*dm12a*md24a
amdb:=am1b*md14b+am2b*md24b+am1b*mm12b*md24b+ad1*dm12b*md24b


adta:=ada+amda
adtb:=adb+amdb

##indirect effect MM1 on dep4
mda:=md12a*dd24a+mm12a*md24a+md14a
mdb:=md12b*dd24b+mm12b*md24b+md14b

## cumulative effect d1 on dep4
dda:=dm12a*md24a+dd12a*dd24a+dd14a
ddb:=dm12b*md24b+dd12b*dd24b+dd14b

## indirect effect of educaiton on dep4
eda:=ed1a*dd12a*dd24a+ed2a*dd24a
edb:=ed1b*dd12b*dd24b+ed2b*dd24b
emda:=ed1a*dm12a*md24a
emdb:=ed1b*dm12b*md24b

edta:=eda+amda
edtb:=edb+emdb

'


res_covAgeEd2_g<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_covAgeEd2_R,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=TRUE
         )
summary(res_covAgeEd2_g,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)

anova(resFre_g,  res_covR_g,res_covAge_g,res_covAgeEd_g,res_covAgeEd2_g)
```
difference is still significant, constraining to make the effect of age being the  same among groups leads to worsening of the fit.

```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_covAgeEd2_g,path=('./Results/LaggedSEM_mi20/gSEX_covAgeEd'))
res$gPartable

```


# testing moderation
```{r}

model_mod<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
Dep_EpiDoc2~Dep_EpiDoc1

MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1

MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1



# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~Age
Dep_EpiDoc1~Age

MM_EpiDoc4~0*Age
MM_EpiDoc2~Age
MM_EpiDoc1~Age


# educação
Dep_EpiDoc4~0*Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4~0*Education
MM_EpiDoc2~Education
MM_EpiDoc1~Education


# interaction terms
#Dep_EpiDoc4~MM2_ed+MM1_ed
#Dep_EpiDoc2~MM1_ed



MM_EpiDoc4~d2_ed+d2_ed
MM_EpiDoc2~d1_ed



MM_EpiDoc4~age_ed
MM_EpiDoc2~age_ed
MM_EpiDoc1~age_ed

# correlation among interaction terms and its original var
#MM2_ed~~Education+MM_EpiDoc2+age_ed+Age+MM1_ed+MM_EpiDoc1+d2_ed+d1_ed
#MM1_ed~~Education+MM_EpiDoc1+age_ed+Age+MM_EpiDoc2+d2_ed+d1_ed
d2_ed~~Education+Dep_EpiDoc2+age_ed+d1_ed+Age+Dep_EpiDoc1
d1_ed~~Education+Dep_EpiDoc1+age_ed+Age+Dep_EpiDoc2
age_ed~~Education+Age
Education~~Age

'


res_mod<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_mod,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=FALSE
         )
summary(res_mod,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_mod,path=('./Results/LaggedSEM_mi20/Moderation'))
res$gPartable

```


# testing moderation with quant
```{r}

model_mod<-'
# evolução
Dep_EpiDoc4~c(dd14a,dd14b)*Dep_EpiDoc1+c(dd24a,dd24b)*Dep_EpiDoc2
Dep_EpiDoc2~c(dd12a,dd12b)*Dep_EpiDoc1

MM_EpiDoc4~c(mm24a,mm24b)*MM_EpiDoc2+c(mm14a,mm14b)*MM_EpiDoc1
MM_EpiDoc2~c(mm12a,mm12b)*MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~c(md24a,md24b)*MM_EpiDoc2+c(md14a,md14b)*MM_EpiDoc1
Dep_EpiDoc2~c(md12a,md12b)*MM_EpiDoc1

MM_EpiDoc4~c(dm24a,dm24b)*Dep_EpiDoc2+c(dm14a,dm14b)*Dep_EpiDoc1
MM_EpiDoc2~c(dm12a,dm12b)*Dep_EpiDoc1


# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~0*Dep_EpiDoc2
MM_EpiDoc4~~0*Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
Dep_EpiDoc2~ad2*Age
Dep_EpiDoc1~ad1*Age

MM_EpiDoc4~c(am4a,am4b)*Age
MM_EpiDoc2~c(am2a,am2b)*Age
MM_EpiDoc1~c(am1a,am1b)*Age


# educação
Dep_EpiDoc4~0*Education
Dep_EpiDoc2~c(ed2a,ed2b)*Education
Dep_EpiDoc1~c(ed1a,ed1b)*Education
MM_EpiDoc4~0*Education
MM_EpiDoc2~0*Education
MM_EpiDoc1~0*Education



# effects on depression =============
##indirect effect age on dep4
ada:=ad1*dd14a+ad2*dd24a+ad1*dd12a*dd24a
adb:=ad1*dd14b+ad2*dd24b+ad1*dd12b*dd24b
amda:=am1a*md14a+am2a*md24a+am1a*mm12a*md24a+ad1*dm12a*md24a
amdb:=am1b*md14b+am2b*md24b+am1b*mm12b*md24b+ad1*dm12b*md24b


adta:=ada+amda
adtb:=adb+amdb

##indirect effect MM1 on dep4
mda:=md12a*dd24a+mm12a*md24a+md14a
mdb:=md12b*dd24b+mm12b*md24b+md14b

## cumulative effect d1 on dep4
dda:=dm12a*md24a+dd12a*dd24a+dd14a
ddb:=dm12b*md24b+dd12b*dd24b+dd14b

## indirect effect of educaiton on dep4
eda:=ed1a*dd12a*dd24a+ed2a*dd24a
edb:=ed1b*dd12b*dd24b+ed2b*dd24b
emda:=ed1a*dm12a*md24a
emdb:=ed1b*dm12b*md24b

edta:=eda+amda
edtb:=edb+emdb


# interaction terms =========
Dep_EpiDoc4~MM2_ed+MM1_ed
Dep_EpiDoc2~MM1_ed



MM_EpiDoc4~d2_ed+d2_ed
MM_EpiDoc2~d1_ed



MM_EpiDoc4~age_ed
MM_EpiDoc2~age_ed
MM_EpiDoc1~age_ed

# correlation among interaction terms and its original var
MM2_ed~~Education+MM_EpiDoc2+age_ed+Age+MM1_ed+MM_EpiDoc1+d2_ed+d1_ed
MM1_ed~~Education+MM_EpiDoc1+age_ed+Age+MM_EpiDoc2+d2_ed+d1_ed
d2_ed~~Education+Dep_EpiDoc2+age_ed+d1_ed+Age+Dep_EpiDoc1
d1_ed~~Education+Dep_EpiDoc1+age_ed+Age+Dep_EpiDoc2
age_ed~~Education+Age
Education~~Age



'


res_mod<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=model_mod,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",# missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group='Sex',
         orthogonal=FALSE
         )
summary(res_mod,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
```{r}

source('./code/parTableToCSV.R')
res<-parTableToCSV(res_mod,path=('./Results/LaggedSEM_mi20/Moderation'))
res$gPartable

```

<!-- # plot -->
## full data
```{r eval=FALSE, include=FALSE}
library(semTools)
source('./code/plotLavaanMI.R')
#lay=get_layout(
 #          '',"Dep_EpiDoc1","", "","", "Dep_EpiDoc4", 
 #          '',"","", "Dep_EpiDoc2","", "", 
 #          "", "",'',"","",'',
 #           "", "",'',"","",'',
 #          '',"","",'MM_EpiDoc2',"","",
 #         '','MM_EpiDoc1',"","","",'MM_EpiDoc4',
 #          "", "",'',"","",'',
 #          "Age", "", "",'',"","",
 #          rows = 7)
lay=get_layout(
  "Age", "","", '',"","",'',"",  "","",
            "", "","", '',"","",'',"",  "","",
   "", "","", '',"","",'Dep_EpiDoc2',"",  "","Dep_EpiDoc4",
          "", '',"", "","", "", "", "","", "", 
            "",'',"","Dep_EpiDoc1", "", "","",  "","", "",
           "Education", "","", '',"","",'',"",  "","",
            "", "","", '',"","",'',"",  "","",
  "",'',"","", "","",'MM_EpiDoc2',"","",  "MM_EpiDoc4",
            "", "","", '',"","",'',"",  "","",
           "",'',"", 'MM_EpiDoc1',"","","","", "", '',
           rows = 10)
plotLavaanMI(model=modelAll0_adj,layout=lay,data.mi=data.mi,estimator='ML',
                       test = "yuan.bentler.mplus",
                       likelihood = "wishart", missing = "FIML",
                       sampling.weights='ipw',
                       orthogonal=TRUE,
                       filepath='./Results/LaggedSEM_mi20/laggedSEM.png')
```


