---
title: "2 Cross-Lagged Structural Equation analysis"
author: "Tomoe Gusberti, Dr. Eng"
date: "2025-03-20"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lavaan)
#library(tidySEM)
#detach("package:lavaan.mi", unload = TRUE)
library(lavaan.mi)
#library(semTools)


library(ggplot2)
```

load data
```{r}
load(file='./Data/Data_Imp20_forSEM.RData')#data.mi
GVAr='Sex'
folderName=paste0('MGroup_',GVAr)
#MGroup_sex
```

# Defining models
## model with (almost) no restriction (almost saturated model)
- we just let DepEpidoc4~agE be 0, based on previous results
```{r}
Model_sat<-'
# evolução
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~Dep_EpiDoc1
MM_EpiDoc2~~Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~Age
Dep_EpiDoc2~Age
MM_EpiDoc2~Age
Dep_EpiDoc1~Age
MM_EpiDoc1~Age

# Escolaridade
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4+MM_EpiDoc2+MM_EpiDoc1~Education
'
```
```{r}
resSatModel<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_sat,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resSatModel,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
```{r}
partable<-parameterEstimates.mi(resSatModel,standardized =TRUE)
partable<-partable%>%mutate(
  label2=paste0(lhs,op,rhs),
  estimate_p=paste0(round(est,3),' (',round(pvalue,3),')'),
  CI=paste0('[',round(ci.lower,3),':',round(ci.upper,3),']')
)%>%select(-c(lhs,op,rhs))
gPartable<-reshape(partable%>%filter(!block==0), idvar = c('label2'),
                                                   timevar = "group", direction = "wide",sep='_')
gPartable%>%select(label2,estimate_p_1,estimate_p_2,CI_1,CI_2)
```


# the extreme test: strong inv
```{r}

resAllEd_g_strongInv<-lavaan.mi::sem.mi(data=data.mi,#[complete.cases(data),],
         model=Model_sat,# missing='pairwise',
         #estimator='MLR', changed for bootstrap
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart", #missing = "FIML", #estimator='WLSMV', later when ordinal variables are added
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE,
         group.equal = c("intercepts", "loadings")
         )


# model comparison tests
lavaan.mi::lavTestLRT.mi(resSatModel,  resAllEd_g_strongInv)

```

# testing partial constrains
## model with restriction on residualCov
```{r}
Model_RCov<-'
# evolução
#Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
Dep_EpiDoc4~Dep_EpiDoc1+Dep_EpiDoc2
MM_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
Dep_EpiDoc2~Dep_EpiDoc1
MM_EpiDoc2~MM_EpiDoc1

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~md1*Dep_EpiDoc1
MM_EpiDoc2~~0* Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~Age
Dep_EpiDoc2~Age
MM_EpiDoc2~Age
Dep_EpiDoc1~Age
MM_EpiDoc1~Age

# Escolaridade
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4+MM_EpiDoc2+MM_EpiDoc1~Education
'
```
```{r}
resRCov<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_RCov,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resRCov,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
### Test significance free model x resRCov
```{r}
lavaan.mi::lavTestLRT.mi(resRCov,resSatModel )
```
## model with restriction on AR components
```{r}
Model_AR<-'
# evolução
#Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
Dep_EpiDoc4~Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~mm14*MM_EpiDoc2+mm04*MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~mm01*MM_EpiDoc1 # freed up considering mi

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+MM_EpiDoc1
MM_EpiDoc4~Dep_EpiDoc2+Dep_EpiDoc1
Dep_EpiDoc2~MM_EpiDoc1
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~0*Dep_EpiDoc1
MM_EpiDoc2~~0* Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4

# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~Age
Dep_EpiDoc2~Age
MM_EpiDoc2~Age
Dep_EpiDoc1~Age
MM_EpiDoc1~Age

# Escolaridade
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4+MM_EpiDoc2+MM_EpiDoc1~Education
'
```
```{r}
resAR<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_AR,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resAR,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
### Test significance free model x AR stricted
```{r}
lavaan.mi::lavTestLRT.mi(resAR,resSatModel )
```

```{r}
partable<-parameterEstimates.mi(resAR,standardized =TRUE)
partable<-partable%>%mutate(
  label2=paste0(lhs,op,rhs),
  estimate_p=paste0(round(est,3),' (',round(pvalue,3),')'),
  CI=paste0('[',round(ci.lower,3),':',round(ci.upper,3),']')
)%>%select(-c(lhs,op,rhs))
gPartable<-reshape(partable%>%filter(!block==0), idvar = c('label2'),
                                                   timevar = "group", direction = "wide",sep='_')
gPartable%>%select(label2,estimate_p_1,estimate_p_2,CI_1,CI_2)
```

## model with restriction on AR and Bi-dir
```{r}
Model_AR_BiD<-'
# evolução
# evolução
#Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
Dep_EpiDoc4~Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~mm14*MM_EpiDoc2+mm04*MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~mm01*MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+0*MM_EpiDoc1
Dep_EpiDoc2~dm21*MM_EpiDoc1
MM_EpiDoc4~0*Dep_EpiDoc2+Dep_EpiDoc1 
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~0*Dep_EpiDoc1
MM_EpiDoc2~~0* Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4


# causalidade - idade
Dep_EpiDoc4~Age
MM_EpiDoc4~Age
Dep_EpiDoc2~Age
MM_EpiDoc2~Age
Dep_EpiDoc1~Age
MM_EpiDoc1~Age

# Escolaridade
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4+MM_EpiDoc2+MM_EpiDoc1~Education

'

```
```{r}
resAR_BiD<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_AR_BiD,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resAR_BiD,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```

### testing if keep beign the same of sat
```{r}
lavaan.mi::lavTestLRT.mi(resAR_BiD,resSatModel )
```
```{r}
lavaan.mi::lavTestLRT.mi(resAR_BiD,resAR,resRCov,resSatModel )
```


```{r}
partable<-parameterEstimates.mi(resAR_BiD,standardized =TRUE)
partable<-partable%>%mutate(
  label2=paste0(lhs,op,rhs),
  estimate_p=paste0(round(est,3),' (',round(pvalue,3),')'),
  CI=paste0('[',round(ci.lower,3),':',round(ci.upper,3),']')
)%>%select(-c(lhs,op,rhs))
gPartable<-reshape(partable%>%filter(!block==0), idvar = c('label2'),
                                                   timevar = "group", direction = "wide",sep='_')
gPartable%>%select(label2,estimate_p_1,estimate_p_2,CI_1,CI_2)
```
## model with restriction on AR, bi-dir (part) and age
```{r}
Model_ARAge<-'
# evolução
#Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
Dep_EpiDoc4~Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~mm14*MM_EpiDoc2+mm04*MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~mm01*MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+0*MM_EpiDoc1
Dep_EpiDoc2~dm21*MM_EpiDoc1
MM_EpiDoc4~0*Dep_EpiDoc2+Dep_EpiDoc1 
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~0*Dep_EpiDoc1
MM_EpiDoc2~~0* Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4


# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~ma4*Age
Dep_EpiDoc2~da2*Age
MM_EpiDoc2~ma2*Age
Dep_EpiDoc1~da1*Age
MM_EpiDoc1~ma1*Age

# Escolaridade
Dep_EpiDoc4~Education
Dep_EpiDoc2~Education
Dep_EpiDoc1~Education
MM_EpiDoc4+MM_EpiDoc2+MM_EpiDoc1~Education

'

```
```{r}
resARAge<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_ARAge,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resARAge,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```
```{r}
lavaan.mi::lavTestLRT.mi(resARAge,resSatModel )
```
```{r}
lavaan.mi::lavTestLRT.mi(resARAge,resAR_BiD,resAR,resRCov,resSatModel )
```


```{r}
partable<-parameterEstimates.mi(resARAge,standardized =TRUE)
partable<-partable%>%mutate(
  label2=paste0(lhs,op,rhs),
  estimate_p=paste0(round(est,3),' (',round(pvalue,3),')'),
  CI=paste0('[',round(ci.lower,3),':',round(ci.upper,3),']')
)%>%select(-c(lhs,op,rhs))
gPartable<-reshape(partable%>%filter(!block==0), idvar = c('label2'),
                                                   timevar = "group", direction = "wide",sep='_')
gPartable%>%select(label2,estimate_p_1,estimate_p_2,CI_1,CI_2)
```

## model with restriction on ED
```{r}
Model_ED<-'
# evolução
#Dep_EpiDoc4~dd03*Dep_EpiDoc1+dd13*Dep_EpiDoc2
Dep_EpiDoc4~Dep_EpiDoc1+dd13*Dep_EpiDoc2
MM_EpiDoc4~mm14*MM_EpiDoc2+mm04*MM_EpiDoc1
Dep_EpiDoc2~dd01*Dep_EpiDoc1
MM_EpiDoc2~mm01*MM_EpiDoc1 

# causalidade - c/ temporalidade/lag
Dep_EpiDoc4~MM_EpiDoc2+0*MM_EpiDoc1
Dep_EpiDoc2~dm21*MM_EpiDoc1
MM_EpiDoc4~0*Dep_EpiDoc2+Dep_EpiDoc1 
MM_EpiDoc2~Dep_EpiDoc1

# corrrelação mesmo t
MM_EpiDoc1~~0*Dep_EpiDoc1
MM_EpiDoc2~~0* Dep_EpiDoc2
MM_EpiDoc4~~Dep_EpiDoc4


# causalidade - idade
Dep_EpiDoc4~0*Age
MM_EpiDoc4~ma4*Age
Dep_EpiDoc2~da2*Age
MM_EpiDoc2~ma2*Age
Dep_EpiDoc1~da1*Age
MM_EpiDoc1~ma1*Age

# Escolaridade
Dep_EpiDoc4~0*Education
Dep_EpiDoc2~de2*Education
Dep_EpiDoc1~Education
MM_EpiDoc4~0*Education
MM_EpiDoc2~0*Education
MM_EpiDoc1~Education

'

```
```{r}
resED<-lavaan.mi::sem.mi(data=data.mi,
         model=Model_ED,
         estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart",
           sampling.weights='ipw',group=GVAr,
         orthogonal=TRUE
         )
summary(resED,fit.meas=TRUE,standardized=TRUE,rsquare=TRUE,ci=TRUE)
```

```{r}
lavaan.mi::lavTestLRT.mi(resED,resSatModel )
```
```{r}
lavaan.mi::lavTestLRT.mi(resED, resARAge,resAR_BiD,resAR,resRCov, resSatModel )
```

## reported on fully adjusted

```{r}

partable<-parameterEstimates.mi(resED,standardized =TRUE)
partable<-partable%>%mutate(
  label2=paste0(lhs,op,rhs),
  estimate_p=paste0(round(est,3),' (',round(pvalue,3),')'),
  CI=paste0('[',round(ci.lower,3),':',round(ci.upper,3),']')
)%>%select(-c(lhs,op,rhs))
gPartable<-reshape(partable%>%filter(!block==0), idvar = c('label2'),
                                                   timevar = "group", direction = "wide",sep='_')
gPartable%>%select(label2,estimate_p_1,estimate_p_2,CI_1,CI_2)
source('./code/parTableToCSV.R')
parTableToCSV(resED,path=paste0('./Results/LaggedSEM_mi20_final/',folderName))

```

# GENERATING PLOTs
## prepare for miPLOt
```{r}
library(mice)
data_bal=complete(data.mi,0)
res_dt0<-sem(data=data_bal,
                   model=Model_ED,
                    estimator = "ML", test = "yuan.bentler.mplus",
         likelihood = "wishart", 
           sampling.weights='ipw',
         orthogonal=TRUE
  )
```

## general plot
```{r}
source('./code/plotLavaanMI.R')
#lay=get_layout(
 #          '',"Dep_EpiDoc1","", "","", "Dep_EpiDoc4", 
 #          '',"","", "Dep_EpiDoc2","", "", 
 #          "", "",'',"","",'',
 #           "", "",'',"","",'',
 #          '',"","",'MM_EpiDoc2',"","",
 #         '','MM_EpiDoc1',"","","",'MM_EpiDoc4',
 #          "", "",'',"","",'',
 #          "Age", "", "",'',"","",
 #          rows = 7)
lay=get_layout(
  "Age", "","", '',"","",'',"",  "","",
            "", "","", '',"","",'',"",  "","",
   "", "","", '',"","",'Dep_EpiDoc2',"",  "","Dep_EpiDoc4",
          "", '',"", "","", "", "", "","", "", 
            "",'',"","Dep_EpiDoc1", "", "","",  "","", "",
           "", "","", '',"","",'',"",  "","",
            "Education", "","", '',"","",'',"",  "","",
  "",'',"","", "","",'MM_EpiDoc2',"","",  "MM_EpiDoc4",
            "", "","", '',"","",'',"",  "","",
           "",'',"", 'MM_EpiDoc1',"","","","", "", '',
           rows = 10)

```

```{r include=FALSE}
library(tidySEM)
parTable<-parameterEstimates.mi(resED, standardized=TRUE)
dfPar<-get_edges(res_dt0)
dfPar<-dfPar%>%mutate(
  path= paste0(lhs,op,rhs)
)
dfPar<-left_join(dfPar%>%select(-c(est,se)),parTable,by=c('lhs','op','rhs'))
dfPar<-dfPar%>%mutate(
  #label=paste(paste(format(std.all,digit=3),'(p=',format(pvalue,digit=3),')'),collapse=';\n')
  pvalue_temp=ifelse(pvalue<0.001,'<0.001',sprintf('=%.3f',pvalue)),
  label=sprintf('%.3f (p%s)',std.all,pvalue_temp)
  )
dfPar$label<-ifelse(dfPar$arrow=='both','',dfPar$label)
```
### automatic
```{r include=FALSE}

nodes <- get_nodes(res_dt0)


ba<-prepare_graph(res_dt0, layout = lay,
                  rect_width = 2.5, # Increase rectangle width
  rect_height = 1, # Increase rectangle height
  )%>%
  #edit_graph( {label = dfPar$rhs})%>%
  edit_graph( {label = dfPar$label},element='edges') %>%
  edit_graph({label_colour = ifelse(dfPar$arrow == "last",ifelse(dfPar$pvalue>0.01,'grey','blue') , "darkgreen")},element='edges')%>%
  #edit_graph({label_colour = ifelse(dfPar$op == "~~",'white',
   #           ifelse(dfPar$arrow == "last",
    #                 ifelse(dfPar$pvalue>0.05,'white','blue') , "darkgreen")                      
     #                              ) },element='edges')%>%
  hide_cov()%>%
  #hide_nonsig()%>%
  #edit_graph({label_colour='blue'},element='nodes')%>%
  #edit_nodes({label_colour=ifelse(grepl('Dep',nodes$name),'blue','darkgrey')    })%>%
  #edit_graph({hide_var = TRUE}) %>% edit_graph({size = c(15, 2)}) %>%
  edit_graph({label_location = 0.35})%>%
  edit_graph({label_size = 3},element='edges')%>%
   edit_graph({label_size = 3},element='nodes')%>%
  #color_nonsig_edges("white") %>%
  color_pos_edges("darkgrey") %>%
  #color_neg_edges("red") %>%
  color_var("white") %>%
  #if_nodes({grepl('Dep',label)},{label_colour='blue'})%>%
  plot(size = c(50, 15))
ba
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/laggedSEM_general.png'), width = 30, height = 15, units = "cm")
```
## groups

```{r}

dfPar<-get_edges(res_dt0)
p <- prepare_graph(res_dt0, layout = lay)
dfNPar<-p$nodes
partable<-parameterEstimates.mi(resED,standardized =TRUE)
dfPar2<-left_join(dfPar%>%select(rhs,lhs,op,arrow,est_sig_std),
                    partable)%>%mutate(
                      pvalue_temp=ifelse(pvalue<0.001,'<0.001',sprintf('=%.3f',pvalue)),
                      label2=sprintf('%.3f (p%s)',std.all,pvalue_temp),
                      label2=ifelse(op=='~~','',label2)
                    )
dfPar2<-dfPar2%>%mutate(
  path= paste0(rhs,op,lhs)
)
dfPar2<-dfPar2%>%group_by(path)%>%mutate(label=paste(paste(group,' ',est_sig_std),collapse=';\n'))
```


```{r echo=TRUE}
b<-prepare_graph(res_dt0, layout = lay)%>%
  #edit_graph( {label = dfPar2$lhs})%>%
  edit_graph( {label = dfPar2$label},element='edges') %>%
  #edit_graph({hide_var = TRUE}) %>% edit_graph({size = c(15, 2)}) %>%
  edit_graph({label_location = 0.3})%>%
  edit_graph({label_size = 2},element='edges')%>%
  color_var("white") %>%
  plot(size = c(15, 2))

```

```{r}
partable<-parameterEstimates.mi(resED,standardized =TRUE)
partable<-partable%>%mutate(
  label= paste0(lhs,op,rhs)
)%>%filter(op=='~')
partable<-partable%>%mutate(
  index=as.numeric(rownames(partable)))

fitms<-fitMeasures(resED)
fits<-paste0('CFI: ',round(fitms[['cfi']],3),' TLI: ',round(fitms[['tli']],3),' RMSEA: ',round(fitms[['rmsea']],3),' SRMR: ',round(fitms[['srmr']],3))
print(fits)
partable$group<-factor(partable$group,labels=c('male','female'))
gall<-ggplot(data=partable,aes(y = factor(group), x = est,color=factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=2)+
  #facet_grid(cols = vars(label))+
  scale_y_discrete(name = "",labels = unique(data_bal$depression)) +theme_bw() +xlab('estimate')+
  theme(strip.text.y = element_text(angle = 0))
gall
library(ggpubr)
ggarrange(b, gall, widths = c(5,3),          ncol = 2)
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/laggedSEM_group1.png'), width = 35, height = 15, units = "cm")
```

```{r include=FALSE}
partable$group<-factor(partable$group,labels=c('male','female'))

partable<-partable%>%mutate(
  rhs_cat=strtrim(rhs,3),
  lhs_cat=strtrim(lhs,3)
)
g<-ggplot(data=partable%>%filter(!rhs=='Age')%>%filter(rhs_cat==lhs_cat),aes(y = factor(group), x = est,color= factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=1)+
  #facet_grid(cols = vars(label))+
  #scale_y_discrete(name = "",labels = unique(data$depression)) +
  theme_bw() +xlab('estimate')+ylab('')+
  theme(strip.text.y = element_text(angle = 0),
        axis.ticks.y=element_blank(),
         axis.text.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.position="bottom",
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.box.margin=margin(-10,0,-10,0),
        panel.margin.y = unit(0, "lines"))+
  labs(color=GVAr,x=NULL)

g_cross_dep<-ggplot(data=partable%>%filter(!rhs=='Age')%>%filter(!rhs_cat==lhs_cat)%>%filter(rhs_cat=='Dep'),aes(y = factor(group), x = est,color= factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=1)+
  #facet_grid(cols = vars(label))+
  #scale_y_discrete(name = "",labels = unique(data$depression)) +
  theme_bw() +xlab('estimate')+ylab('')+
  theme(strip.text.y = element_text(angle = 0),
        axis.ticks.y=element_blank(),
         axis.text.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.position="bottom",
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.box.margin=margin(-10,0,-10,0),
        panel.margin.y = unit(0, "lines"))+
  labs(color=GVAr,x=NULL)
g_cross_Mmorb<-ggplot(data=partable%>%filter(!rhs=='Age')%>%filter(!rhs_cat==lhs_cat)%>%filter(rhs_cat=='MM_'),aes(y = factor(group), x = est,color= factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=1)+
  #facet_grid(cols = vars(label))+
  #scale_y_discrete(name = "",labels = unique(data$depression)) +
  theme_bw() +xlab('estimate')+ylab('')+
  theme(strip.text.y = element_text(angle = 0),
        axis.ticks.y=element_blank(),
         axis.text.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.position="bottom",
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.box.margin=margin(-10,0,-10,0),
        panel.spacing.y = unit(0, "lines"))+
  labs(color=GVAr,x=NULL)
g_cross_Mmorb
```

```{r include=FALSE}
g_Idade<-ggplot(data=partable%>%filter(rhs=='Age'),aes(y = factor(group), x = est,color= factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=1)+
  #facet_grid(cols = vars(label))+
  #scale_y_discrete(name = "",labels = unique(data$depression)) +
  theme_bw() +xlab('estimate')+ylab('')+
  theme(strip.text.y = element_text(angle = 0),
        axis.ticks.y=element_blank(),
         axis.text.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.position="bottom",
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.box.margin=margin(-10,0,-10,0),
        panel.margin.y = unit(0, "lines"))+
  labs(color='Depression',x=NULL)
g_Idade
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/LaggedSEM_group_forestplot.png'), width = 5, height = 15, units = "cm")
```
```{r include=FALSE}
g_ED<-ggplot(data=partable%>%filter(rhs=='Education'),aes(y = factor(group), x = est,color= factor(group)))+
  geom_point(shape = 18) +  
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper), height = 0.25) +
  geom_vline(xintercept =0, color = "red",  cex = 0.5, alpha = 0.5)+
    facet_wrap( ~ label,ncol=1)+
  #facet_grid(cols = vars(label))+
  #scale_y_discrete(name = "",labels = unique(data$depression)) +
  theme_bw() +xlab('estimate')+ylab('')+
  theme(strip.text.y = element_text(angle = 0),
        axis.ticks.y=element_blank(),
         axis.text.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.position="bottom",
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.box.margin=margin(-10,0,-10,0),
        panel.margin.y = unit(0, "lines"))+
  labs(color='Depression',x=NULL)
g_Idade
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/LaggedSEM_group_forestplot.png'), width = 5, height = 15, units = "cm")
```

## general plot + forest plot for groups
```{r include=FALSE}
ggarrange(ba+theme(plot.margin = unit(c(-0.9,-0.9,-0.9,-0.9), 'lines')), 
          ggarrange(g+theme(plot.margin = unit(c(1,0,0,-0.9), 'lines'))+
            theme(strip.background =element_rect(fill="white",colour=NA),
                  panel.border = element_blank())+ theme(axis.text.x = element_text(size=8)),
  ggarrange(g_cross_Mmorb+theme(plot.margin = unit(c(1,0,0,-0.9), 'lines'))+
              theme(strip.background =element_rect(fill="white",colour=NA),
                    panel.border = element_blank())+ theme(axis.text.x = element_text(size=8)),
            g_cross_dep+theme(plot.margin = unit(c(0,0,0,-0.9), 'lines'))+
              theme(strip.background =element_rect(fill="white",colour=NA),
                    panel.border = element_blank())+ theme(axis.text.x = element_text(size=8)), 
            legend='none', heights =  c(3,5),nrow = 2), 
  ggarrange(g_Idade+theme(plot.margin = unit(c(1,0.1,0,-0.9), 'lines'))+
              theme(strip.background =element_rect(fill="white",colour=NA),
                    panel.border = element_blank())+ theme(axis.text.x = element_text(size=8)),
            g_ED+theme(plot.margin = unit(c(0,0,0,-0.9), 'lines'))+
              theme(strip.background =element_rect(fill="white",colour=NA),
                    panel.border = element_blank())+ theme(axis.text.x = element_text(size=8)), 
            legend='none', heights =  c(5,3),nrow = 2)+
  theme(strip.background =element_rect(fill="white",colour=NA), 
        panel.border = element_blank()), 
                                          common.legend = TRUE,widths = c(2,2,2),
          ncol = 3), widths = c(6,3),
          ncol = 2)
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/laggedSEM_group_and_general.png'), width = 30, height = 15, units = "cm")
```

## just ed and age
```{r include=FALSE}
ggarrange(ba+theme(plot.margin = unit(c(-0.9,-0.9,-0.9,-0.9), 'lines')), 
          ggarrange(
            ggarrange(g_cross_Mmorb+theme(plot.margin = unit(c(1,0,0,-0.9), 'lines'))+
                        theme(strip.background =element_rect(fill="white",colour=NA),
                              panel.border = element_blank())+ 
                        theme(axis.text.x = element_text(size=6)),
                      g_cross_dep+theme(plot.margin = unit(c(0,0,0,-0.9), 'lines'))+
                        theme(strip.background =element_rect(fill="white",colour=NA),
                              panel.border = element_blank())+ 
                        theme(axis.text.x = element_text(size=6)), 
                      legend='none', heights =  c(5,5),nrow = 2), 
            g_Idade+theme(plot.margin = unit(c(1,0.1,0,-0.9), 'lines'))+
                        theme(strip.background =element_rect(fill="white",colour=NA),
                              panel.border = element_blank())+ 
                        theme(axis.text.x = element_text(size=6)),
          g_ED+theme(plot.margin = unit(c(1,1,0,-0.9),'lines'))+
                        theme(strip.background =element_rect(fill="white",colour=NA),
                              panel.border = element_blank())+ 
                        theme(axis.text.x = element_text(size=6)), 
          common.legend = TRUE,widths = c(2,2,3), ncol = 3), 
widths = c(6,3),          ncol = 2)
ggsave(filename=paste0('./Results/LaggedSEM_mi20_final/',folderName,'/laggedSEM_group_and_general2.png'), width = 30, height = 15, units = "cm")
```


